---
title: "CodeBook"
author: "R. Torres"
date: "7/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting and Cleaning Data Course Project

This documents presents the information related to the run_analysis program to collect, transform and generate data. The following steps describe the code sequence and outputs.

  **a. Load all data into tables**
  
Dataset downloaded from [link](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip) and stored locally in a defined path.
Setwd must be used to direct R to the designated location
  
##### Data Frames:
```{r, chunk-code, include=FALSE}
library(dplyr)
features <- read.table("features.txt", col.names = c("n","feature"))
train_x <- read.table("train/X_train.txt", col.names = features$feature)
train_y <- read.table("train/y_train.txt", col.names = "code")
test_x <- read.table("test/X_test.txt", col.names = features$feature)
test_y <- read.table("test/y_test.txt", col.names = "code")
train_subject <- read.table("train/subject_train.txt", col.names = "subject")
test_subject <- read.table("test/subject_test.txt", col.names = "subject")
act_lbls <- read.table("activity_labels.txt", col.names = c("code", "activity"))

#Data frame to use as merged
merged <- tibble()
#join test and train data for x and y
x_data <- rbind(train_x, test_x)
y_data <- rbind(train_y, test_y)

sbj <- rbind(train_subject, test_subject)

#Merge all info together
merged <- cbind(sbj, x_data, y_data)

```
`Features`, total rows: **`r nrow(features)`** . Contains accelerometer and gyroscope data\
`train_x`, total rows: **`r nrow(train_x)`**. Training data\
`train_y`, total rows: **`r nrow(train_y)`**. Training activity labels\
`test_x`, total rows: **`r nrow(test_x)`**. Test data\
`test_y`, total rows: **`r nrow(test_y)`**. Test activity labels\
`train_subject`, total rows: **`r nrow(train_subject)`**. Contains the data of training volunteers\
`test_subject`, total rows: **`r nrow(test_subject)`**. Contains the data of test volunteers\
`act_lbls`, total rows: **`r nrow(act_lbls)`**. List of activities performed by subjects and the codes associated to them.
\

**b. Merges the training and the test sets to create one data set.**

- `x_data` <- Generated after merging train_x and test_x using **rbind()**
- `y_data` <- Generated after merging train_y and test_y using **rbind()**
- `sbj` <- Generated after merging train_subj and test_subj using **rbind()**
- `merged` <- Data frame containing training, test and subject data together using **cbind()**

Total rows and column after merging all data: **`r nrow(merged)`** rows and **`r ncol(merged)`** columns

**c. Extracts only the measurements on the mean and standard deviation for each measurement.**


`TidyData`: Generated filtering mergedData using mutate function `%>%`. Only measurements containing `mean` or `standard deviation` are used. Columns `Code` and `Subject` are selected

**d. Uses descriptive activity names to name the activities in the data set**

Using the numbers of `code`, the names of the activities are taken from `act_lbls` as follows:

- `Code` 1: `WALKING`
- `Code` 2: `WALKING_UPSTAIRS`
- `Code` 3: `WALKING_DOWNSTAIRS`
- `Code` 4: `SITTING`
- `Code` 5: `STANDING`
- `Code` 6: `LAYING`
  
**e. Appropriately labels the data set with descriptive variable names.**

With instruction `gsub`, unclear names are changed for more descriptive variables:

`Acc`, Replaced by Accelerometer\
`Gyro`, Replaced by Gyroscope\
`angle`, Replaced by Angle\
`gravity`, Replaced by Gravity\
`tBody`, Replaced by TimeBody\
`mean`, Replaced by Mean\
`std`, Replaced by STD\
`^t`, Replaced by Time\
`^f`, Replaced by Frecuency\

**f. Create a data set with the average of each variable for each activity and each subject**

- Using `mutate, group_by and summarise_all` functions, each variable is combined and the average is calculated using `mean`.
- Data is exported to file `TidyDataOutput.csv` for easy reading and understanding.



